---
title: "EDLD652_Lab_5"
author: "Michelle Cui"
date: "2025-02-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r include=FALSE}
library(tidyverse)
library(here)
library(scales)
library(ggforce)
library(ggrepel)
library(gghighlight)
theme_set(theme_minimal(base_family = "Times"))
```

```{r}
gap <- read_csv(here("data/gapminder.csv"))
```

### Lab Tasks

1. Recreate the final plot you generated in Lab PS-2 with just the countries in the American Continent

```{r}

gap_1987_am <- gap %>% 
  filter(year == "1987", continent == "Americas")

gap_1987_am %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp,color = continent, shape = continent, size = pop)) +
  geom_point()+
  scale_color_viridis_d()+
  scale_shape_manual(values = c(5))+
  scale_size_continuous(labels = label_comma())+
  scale_x_log10(labels = label_dollar())+
  labs(title = "Life Expectancy at Birth in the America Countries over GDP",
       x = "GDP per Capita (dollars)",
       y = "Life Expectancy (years)")

```

2. Add labels to the countries as there are fewer categories to compare

```{r, fig.width=8, fig.height=6, dpi=300}

gap_1987_am %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp,color = continent, shape = continent, size = pop)) +
  geom_point()+
  scale_color_viridis_d()+
  scale_shape_manual(values = c(5))+
  scale_size_continuous(labels = label_comma())+
  scale_x_log10(labels = label_dollar())+
  geom_text(aes(label = country),
            size = 3, alpha = 0.5, hjust = 1.2, vjust = 0) + labs(title = "Life Expectancy at Birth in the America Countries over GDP",
       x = "GDP per Capita (dollars)",
       y = "Life Expectancy (years)")
```

3. Highlight (hint: use gghighlight) only those countries that have life expectancy greater than 75.

```{r}
gap_1987_am %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop)) +
  geom_point() +
  scale_color_viridis_d() +
  scale_shape_manual(values = c(2))+
  scale_size_continuous(labels = label_comma())+
  scale_x_log10(labels = label_dollar()) +
  gghighlight(lifeExp > 75) +
  geom_text_repel(data = filter(gap_1987_am, lifeExp > 75),
                  aes(label = country))+
  labs(title = "Life Expectancy at Birth in the America Countries over GDP",
       x = "GDP per Capita (dollars)",
       y = "Life Expectancy (years)")
  
```

4. Try a marked up elipse(hint: use ggforce library) and geom_mark_elipse) to highlight the two outliers on the SW grid of the plot instead of gghighlight

```{r}
gap_1987_am %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop)) +
  geom_point(aes(color = continent)) +
  scale_color_viridis_d()+
  scale_shape_manual(values = c(2))+
  scale_size_continuous(labels = label_comma())+
  scale_x_log10(labels = label_dollar()) +
  geom_mark_ellipse(
    data = filter(gap_1987_am, lifeExp < 60 & gdpPercap < 3000),
    aes(x = gdpPercap, y = lifeExp))+
  labs(title = "Life Expectancy at Birth in the America Countries over GDP",
       x = "GDP per Capita (dollars)",
       y = "Life Expectancy (years)")
```

### Which one do you prefer between #3 and #4?

Answer: I like plot #3 more than #4 because the visualization on #3 is much clear and neat. It presents the outliers with a simple and direct approach. It might also because the outliers are not large in the dataset. 
