---
title: "EDLD652_Lab_PS3"
author: "Michelle Cui"
date: "2025-02-12"
output:
  html_document:
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cho = TRUE,
                      eval = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.width = 8,
                      fig.height = 6)
```

```{r include=FALSE}
library(tidyverse)
library(here)
library(scales)
library(ggforce)
library(ggrepel)
library(gghighlight)
library(ggplot2)
library(dplyr)
library(forcats)
library(maps)
library(scales)
library(colorblindr)
library(patchwork)
theme_set(theme_minimal(base_family = "Times"))
```

**Data**

```{r}
honey <- read.csv(here("data","honeyproduction.csv"))
#summary(honey)
#str(honey)
```

**Lab Tasks**

1. Create two plots from the same data. Choose wisely - so that they are two different types of plots (also note: plots with facet_wrap doesn’t count here!)
```{r}
honey_98_12 <- honey %>% 
  filter(year == 1998 | year == 2012) %>% 
  select(state, year, stocks) %>% 
  mutate(state = fct_reorder(state, stocks))

p1 <- ggplot(honey_98_12, aes(x = state, y = stocks, fill = year)) +
  geom_col(position = "dodge") +
  labs(title = "Honey Stocks by State in 1998 vs 2012",
       x = "State",
       y = "Honey Stocks",
       fill = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 5))+
  scale_y_continuous(labels = label_comma())


honey_2012 <- honey %>% 
  filter(year == 2012)

honey_2012 <- honey_2012 %>% 
  mutate(state = tolower(state.name[match(state, state.abb)]))

us_states <- map_data("state")

map_data_merged <- us_states %>% 
  left_join(honey_2012, by = c("region" = "state"))

p2 <- ggplot(map_data_merged, aes(x = long, y = lat, group = group, fill = prodvalue)) +
  geom_polygon(color = "white") +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  scale_fill_continuous(name = "Production Value", labels = label_dollar()) +
  labs(title = "Honey Production Value by State(2012)",
       x = "Longitude",
       y = "Latitude",
       fill = "Production Value ($)")


```

2. Now combine those plots side-by-side using library(patchwork) we learned in week 5

```{r}
p1 + p2
```

3. Now combine those plots by stacking them vertically (also using library(patchwork) we learned in week 5

```{r}
p1/p2
```

4. Which one do you prefer when comparing #1 vs. #2 vs. #3? Reflect briefly.

Answer: I like the #3 most because the side by side stacking makes the plots look clustered. However, the #3 is not satisfying because the width of the p2 is affected by p1. I wonder how can I fix it. So I tried a few things.

```{r}
p1/p2 + plot_layout(heights = c(1,0.7)) # Equal heights. This doesn't work well

p1/p2 + plot_layout(guides = "collect", heights = c(1,1))
# This one looks slightly better

(p1 + plot_spacer()) / p2 + plot_layout(heights = c(1,1))
```

